---
title: "Take-home Exercise 6"
description: |
  Explore and reveal the patterns of community interactions of the city of Engagement, Ohio USA by using social network analysis approach.
author:
  - name: Huan Li
    url: https://linkedin.com/in/huan-li-ab7498124/
    affiliation: SMU, SCIS, Master of IT in Business
    affiliation_url: https://scis.smu.edu.sg/master-it-business/about-mitb-main
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# 1. Overview

Reveal the patterns of community interactions of the city of Engagement, Ohio USA by using social network analysis approach.




# 2. Data Preparation

## 2.1 Installing and loading the required libraries

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

The chunk code on the right will do the trick.
```{r}

```



## 2.2 Importing the dataset

The code chunk below imports *participants.csv* and *FinancialJournal.csv* into R environment using [**read_csv()**](https://readr.tidyverse.org/reference/read_delim.html) function of [**readr**](https://readr.tidyverse.org/) package.

```{r}


```


It is always a good practice to examine the imported data frame before further analysis is performed.

For example, *kable()* can be used to review the structure of the imported data frame.

Let's take an overview of the datasets


```{r}

```


## 2.3 Data Wrangling



# 3. Visulisations and Insights