---
title: "Take-home Exercise 1"
description: |
  A new article created using the Distill format.
author:
  - name: Huan Li 
    url: https://linkedin.com/in/huan-li-ab7498124/

date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Overview

This take-home exercise aims to sharpen the skill of building data visualisation programmatically using ggplot2

# Installing and loading the required libraries

The chunk code below will do the trick.

```{r}
packages = c('tidyverse')

for (p in packages){
  if(!require(p, character.only = T)){
    install.package(p)
  }
  library(p, character.only = T)
}
```
# Importing Data

The code chunk below import *Participants.csv* from data folder into R by using [`read_csv()`](https://readr.tidyverse.org/reference/read_delim.html) of [**readr** package](https://readr.tidyverse.org/index.html) and save it as an tibble data frame called *participants*. 

```{r}
participants <- read_csv("data/Participants.csv")
```
# Analysis

The code chunk below plot a bar chart by using [`geom_bar()`](https://ggplot2.tidyverse.org/reference/geom_bar.html) by [**ggplot2 package**](https://ggplot2.tidyverse.org/index.html).

# Distribution of features

There are 7 different columns in this data set, but participant Id column which presents different participant, has over 1000 unique values. I will not include this feature in my analysis since is not a meaningful feature for participants. As for the rest features, plot the distribution for each one as a preparation for further analysis. 

```{r}
library(patchwork)

p1 <- ggplot(data=participants, 
             aes(x = joviality)) +
  geom_histogram(bins = 10,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,1.2)) +
  ggtitle("Distribution of joviality")

p2 <- ggplot(data=participants, 
             aes(x = age)) +
  geom_histogram(bins = 10, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(16, 70)) +
  ggtitle("Distribution of Age")

p1 + p2
```

# Distribution of age and joviality

# Distribution of joviality by No.of households
