---
title: "Take-home Exercise 6"
description: |
  Explore and reveal the patterns of community interactions of the city of Engagement, Ohio USA by using social network analysis approach.
author:
  - name: Huan Li
    url: https://linkedin.com/in/huan-li-ab7498124/
    affiliation: SMU, SCIS, Master of IT in Business
    affiliation_url: https://scis.smu.edu.sg/master-it-business/about-mitb-main
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# 1. Overview

In this project, we need to reveal the patterns of community interactions of the city of Engagement, Ohio USA by using social network analysis approach.

Processes include:
  - create graph object data frames, manipulate them using appropriate functions of dplyr, lubridate, and tidygraph,
  - build network graph visualisation using appropriate functions of ggraph,
  - compute network geometrics using tidygraph,
  - build advanced graph visualisation by incorporating the network geometrics, and
  - build interactive network visualisation using visNetwork package.

# 2. Data Preparation

## 2.1 Installing and loading the required libraries

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

Here four network data modelling and visualisation packages will be installed and launched. They are igraph, tidygraph, ggraph and visNetwork. Beside these four packages, tidyverse and lubridate, an R package specially designed to handle and wrangling time data will be installed and launched too.

The chunk code on the right will do the trick.

```{r}
packages = c('igraph', 'tidygraph', 
             'ggraph', 'visNetwork', 
             'lubridate', 'clock',
             'tidyverse', 'graphlayouts')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```



## 2.2 Importing the dataset

There are two data sets needed. One contains the nodes data and the other contains the edges (also know as link) data.

The edges data

```{r}


```


It is always a good practice to examine the imported data frame before further analysis is performed.

For example, *kable()* can be used to review the structure of the imported data frame.

Let's take an overview of the datasets


```{r}

```


## 2.3 Data Wrangling



# 3. Visulisations and Insights