---
title: "Take-home Exercise 5"
description: |
  Visualising and analysing social areas and locations with traffic bottleneck of the city of Engagement, Ohio USA.

author:
  - name: Huan Li
    url: https://linkedin.com/in/huan-li-ab7498124/
    affiliation: SMU, SCIS, Master of IT in Business
    affiliation_url: https://scis.smu.edu.sg/master-it-business/about-mitb-main
# date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_float: true
#    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





# 1. Overview

Based on dataset [*VAST Challenge 2022*](https://vast-challenge.github.io/2022/), we will explore  and characterize the distinct areas of the city , and characterize the travel patterns to identify potential bottlenecks or hazards, and examine how these patterns change over time. The operation was carried out on Rstudio and main packages used are [**sf**](https://r-spatial.github.io/sf/), [**tmap**](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html) and [**tidyverse**](https://www.tidyverse.org/). 

Questions to be addressed are:

- Social areas of the city of Engagement, Ohio USA.
- Visualising and analysing locations with traffic bottleneck of the city of Engagement, Ohio USA.

# 2. Data Preparation

## 2.1 Installing libraries

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

sf IS an R package specially designed to handle geospatial data in simple feature objects.

The chunk code on the right will do the trick.

```{r}
packages = c('sf','tmap','tidyverse','clock',
             'lubridate','sftime','rmarkdown')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}
```



## 2.2 Importing wkt data

Well-known text (WKT) is a human readable representation for spatial objects like points, lines, or enclosed areas on a map.

Import geospatial data in wkt format into R and saved the imported data as simple feature objects by using sf package

In the code chunk below, read_sf() of sf package is used to parse School.csv Pubs.csv, Apartments.csv, Buildings.csv, Employer.csv, and Restaurants.csv into R as sf data.frames.

```{r}
schools <- read_sf("data/Schools.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")

apartments <- read_sf("data/Apartments.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")

buildings <- read_sf("data/Buildings.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")

employers <- read_sf("data/Employers.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")

jobs <- read_sf("data/Jobs.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")

participants <- read_sf("data/Participants.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")

pubs <- read_sf("data/Pubs.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")

restaurants <- read_sf("data/Restaurants.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")
```


```{r,echo=FALSE, eval=FALSE}
logs <- read_sf("data/wkt/ParticipantStatusLogs1.csv", 
                options = "GEOM_POSSIBLE_NAMES=currentLocation")
```


It is always a good practice to examine the imported data frame before further analysis is performed.

Let's take an overview of the datasets


```{r}
print(buildings)
print(apartments)
```


## 2.3 Data Wrangling

```{r, echo=FALSE, eval=FALSE}
logs_selected <- logs %>%
  mutate(Timestamp = clock::date_time_parse(timestamp,
                                     zone = "",
                                     format = "%Y-%m-%dT%H:%M:%S")) %>%
  mutate(day = clock::get_day(Timestamp)) %>%
  filter(currentMode =="Transport")
```

```{r, echo=FALSE, eval=FALSE}
write_rds(logs_selected,
          "data/rds/logs_selected.rds")
```

```{r}
logs_selected <- read_rds("data/rds/logs_selected.rds")
```

```{r}
print(logs_selected)
```


# 3. Visulisations and Insights

## 3.1 Distinct Social Areas

Characterize the distinct social areas of the [**city of Engagement, Ohio USA**](https://vast-challenge.github.io/2022/).

### 3.1.1 Building Types Map

```{r}
buildingType <- tm_shape(buildings)+
tm_polygons(col = "buildingType",
           palette="Accent",
           border.col = "black",
           border.alpha = .5,
           border.lwd = 0.5)+
tm_layout(main.title = "Building Types Map",
          main.title.position = "center",
          main.title.size = 1,
          frame = FALSE)+
tm_compass(size = 2,
           position = c('right', 'top'))

buildingType
```

**Insights**

- Figure above shows that there are mainly three areas for commercial use - each have one in the middle, the north and the south, and respectively surrounded by the residential areas.

- We can tell that there are mainly two large school zones and respectively in the north and the south, and also two small school zone in the west.

### 3.1.2 Facility Map

```{r}
label <- c('Restaurant', 'Pub', 'Employer', 'Apartment', 'School')
color <- c('blue', 'green', "red", 'purple', 'yellow')

facilitiesMap <- tm_shape(buildings)+
tm_polygons(col = "grey60",
           size = 1,
           border.col = "black",
           border.lwd = 1) +
tm_shape(pubs) +
  tm_dots(col = "green", size = 0.3, alpha= 0.8) +
tm_shape(restaurants) +
  tm_dots(col = "blue", size = 0.3, alpha= 0.8) +
tm_shape(schools) +
  tm_dots(col = "yellow", size = 0.3, alpha= 0.8)+
tm_shape(employers) +
  tm_dots(col = "red") +
tm_shape(apartments) +
  tm_dots(col = "purple") +
tm_add_legend(title = 'Facilities',
              type = 'symbol',
              border.col = NA,
              labels = label,
              col = color) +
tm_layout(main.title = 'Facilities Map of Engagemnt City, Ohio USA',
          main.title.size = 1,
          frame = FALSE) +
tm_compass(size = 2,
           position = c('right', 'top'))+
tm_credits('Source: VAST Challenge 2022')

facilitiesMap
```

**Insights**

- The map above gives us more details of the city's facilities' layout.There are more restaurants and pubs in the middle and northwest of the city, so we can deduct that there might be more traffic in the weekends in those areas.

- Compared with the southeast corner, there are more apartments in the northwest, so might cause more traffic in the weekdays.

```{r}
tmap_arrange(buildingType, facilitiesMap, widths = c(1))
```

**Insights**

- The integrated figure shows that those restaurants, pubs and employers are located in the commercial areas that mainly in the middle.

- The residential zone are at the edges of the city, and the norther , the more residents density.


## 3.2 Traffic Situation

Where are the busiest areas in Engagement? Are there traffic bottlenecks that should be addressed? 

### 3.2.1 General Traffic Situation

#### Computing the haxegons

In the code chunk below, st_make_grid() of sf package is used to create haxegons

```{r}
hex <- st_make_grid(buildings, 
                    cellsize=100, 
                    square=FALSE) %>%
  st_sf() %>%
  rowid_to_column('hex_id')
plot(hex)
```

#### Performing point in polygon count

In the code chunk below, st_join() of sf package is used to count the number of event points in the hexagons.

```{r}
points_in_hex <- st_join(logs_selected, 
                        hex, 
                        join=st_within) %>%
  st_set_geometry(NULL) %>%
  count(name='pointCount', hex_id)
head(points_in_hex)
```

#### Performing relational join

In the code chunk below, left_join() of dplyr package is used to perform a left-join by using hex as the target table and points_in_hex as the join table. The join ID is hex_id.

```{r}
hex_combined <- hex %>%
  left_join(points_in_hex, 
            by = 'hex_id') %>%
  replace(is.na(.), 0)
```

#### Plotting the hexagon binning mapp

In the code chunk below, tmap package is used to create the hexagon binning map.

```{r}
traffic <- tm_shape(hex_combined %>%
                      filter(pointCount > 0))+
  tm_fill("pointCount",
          n = 8,
          style = "quantile") +
  tm_borders(alpha = 0.1)+
  tm_layout(main.title = 'Traffic of Engagemnt City, Ohio USA',
            main.title.size = 1,
            frame = FALSE)
traffic
```

```{r}
tmap_arrange(facilitiesMap,traffic, widths = c(1))
```

**Insights**

- Map above shows that the main routes connecting the the north-west to the south area and wet to the east of the city are likely to see more traffic.

- There is more traffic in places having schools, restaurant, pubs.





# 4. Conclusion

Well-known text (WKT) is a human readable representation for spatial objects like points, lines, or enclosed areas on a map, and helps when doing geo-spatial visualizations,

During this nexercise, we learned how to import geospatial data in wkt format into R and saved the imported data as simple feature objects by using sf package, to map geospatial data using tmap package, to process movement data by using sf and tidyverse packages,and to visualise movement data by using tmap and ggplot2 package.
